---
title: 'CS410 Group Project - foodcasting R code'
author: "Fall 2019, by James Banasiak (jamesmb3) Mark Hornback (markh4) "
date: 'Due: Sunday, Dec 15 by 11:59 PM Pacific Time'
geometry: margin=.1in
fontsize: 11pt
output:
  pdf_document:
    toc: yes
    toc_depth: 2
    df_print: kable
    pandoc_args: [
      "-V", "classoption=twocolumn"
    ]
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
header-includes:
  - \usepackage[ruled,vlined,linesnumbered]{algorithm2e}
  - \usepackage[fontsize=10pt]{scrextend}
---


```{r setup, include=FALSE}
  knitr::opts_chunk$set(include = TRUE)  # TRUE for solution; FALSE for questions set

  knitr::opts_chunk$set(echo = TRUE)
  knitr::opts_chunk$set(message = FALSE)
  knitr::opts_chunk$set(warning = FALSE)
  knitr::opts_chunk$set(fig.height = 6, fig.width = 8, out.width = '50%', fig.align = "center")
  options(width = 90)
```

```{css, echo=FALSE}
.solution {
background-color: #e6ffe6;
}
```


```{r echo=F}
# a better preinstaller - you may need to run to complile from source
########################################################################
# OSX Notes
########################################################################
# osx install xcode and run
#     xcode-select --install
#     brew cask install gfortran
#
REQUIRED_LIBS<-c('knitr','kableExtra','rpart','tidyverse','tidytext','widyr','glmnet','doMC','ggplot2','stringr','dplyr','wordcloud','e1071','tm','caret','quanteda','tidytext','widyr','doParallel','text2vec','RSQLite','dplyr')

# automatically install any missing packages - set
if (length(setdiff(REQUIRED_LIBS, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(REQUIRED_LIBS, rownames(installed.packages())),repos = "http://cran.us.r-project.org")
}
# require the libraries mentioned in REQUIRED_LIBS, hiding the non warning output
invisible(lapply(REQUIRED_LIBS,require,character.only=TRUE))
```


# Project Description

We set off on a journey to predict ratings of restaurants from menu item data.

# Gathering the corpus

The description of the elements contained within the data set is:


For the business listings -
- `slug` - a unique portion of the url, typically dashed business name for http://{host}.com/path/{slug}
- `categories` -  the type of foods the restaurant serves, eg Pizza, Chinese, Mexican
- `distance`- the distance from the centroid of the zip code
- `name` - the name of the restaurant
- `price_level` - a yelp categorical price level \$= under \$10. \$\$=11-30. \$\$\$=31-60. \$\$\$\$= over 61 USD
- `rating`- **predictor**   the yelp rating associated with a restaurnat
- `review_count`- the number of reviews that were taken to establish the rating
- `url` - the url from yelp
- `lat` -the latitude of the location of the restaurant
- `lng` - the longitude of the location of the restaurant
- `Sp1` - a spacer (not used)
- `type` - the type - all `natural` words
- `homeurl` - the path portion of the url
- `resource_id1` - a resource id used in yelp specific api
- `resource_id2` - a resource id used in yelp specific api
- `lat2` - the latitude again resulting from a join within scraper
- `lng2` -  the longitude again resulting from a join within scraper

For the menuitems-
- `slug` -  a unique portion of the url, typically dashed business name for http://{host}.com/path/{slug}
- `title` - a menuitem title eg `Chicken Caesar Salad`
- `description` -  the longer description of the title eg `Grilled chicken, romaine, Parmesan, tomatoes and Caesar dressing`
- `price` -  the price of the menuitem eg `7.99`


# Data processing

```{r}
# set cpu cores
CPU_CORES <- parallel::detectCores()

# setwd("~/git/foodcasting")
conn <- dbConnect(RSQLite::SQLite(), "sql-lite-cache/foodcasting.db")
# get main corpus
df_summary <- dbGetQuery(conn,"SELECT * from summary t1 WHERE t1.slug is not null and t1.slug<>'slug'")
df_menu <- dbGetQuery(conn,"SELECT * from menu t1 WHERE t1.slug is not null and t1.slug<>'slug'")
df <- dbGetQuery(conn,"SELECT t1.slug,t1.categories,t1.rating,t1.review_count,t1.lat,t1.lng,t2.title,t2.description,t2.price from summary t1 left join menu t2 on (t1.slug=t2.slug) WHERE t1.slug is not null and t1.slug<>'slug'")

# remove null  price  and over 5K menuitems
df <- df %>% filter(!is.na(price))  %>% filter(price <= 5000)  %>% filter(!is.na(rating))
# show structure
str(df)
# price prob best in log scale
par(mfrow=c(1,2), mar = rep(4, 4))
hist(log(df$price),main="Histogram of log(price)")
hist(log(df$rating),main="Histogram of log(rating)")
# consolidate chr ? too many


```

```{r}
# plot basic price vs rating to establish a correlation
ggplot(df, aes(price, rating)) +
  geom_point(alpha = .1) +
  geom_smooth(method = "lm") +
  scale_x_log10()

x<-log(df$price)
x[which(!is.finite(x))]<-NA

# show simple model, there is something here worth exploring
model1 <- lm(x ~ review_count + rating + lat + lng, data = df)
summary(model1)

```

There does appear to be signficance of pr

```{r}

# start by expanding words remove stop words
df_title <- df %>%
  mutate(id = row_number()) %>%
  unnest_tokens(word, title) %>%
  anti_join(stop_words, by = "word") %>%
  filter(!word %in% c("ls"),
         str_detect(word, "[a-z]"))
# lets look at top 30 words inside title

```
```{r}

df_title %>%
  count(word, sort = TRUE) %>%
  head(20) %>%
  mutate(word = fct_reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_col() +
  coord_flip()
```

```{r}

df_desc <- df %>%
  mutate(id = row_number()) %>%
  unnest_tokens(word, description) %>%
  anti_join(stop_words, by = "word") %>%
  filter(!word %in% c("ls"),
         str_detect(word, "[a-z]"))
# lets look at top 30 words inside desc
```
```{r}
df_desc %>%
  count(word, sort = TRUE) %>%
  head(20) %>%
  mutate(word = fct_reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_col() +
  coord_flip()

```



```{r}
df_title_agg <- df_title %>%
  distinct(id, word) %>%
  add_count(word) %>%
  filter(n >= 100)

df_desc_agg <- df_desc %>%
  distinct(id, word) %>%
  add_count(word) %>%
  filter(n >= 100)


```

```{r}
# correlation of pairs of items
# df_title_agg %>%
#   pairwise_cor(word, id, sort = TRUE)
#
#
# df_desc_agg %>%
#   pairwise_cor(word, id, sort = TRUE)

# create a sparse matrix from row names, column names, and values in a table
m_title <- df_title_agg %>% cast_sparse(id, word)
m_desc <- df_desc_agg %>% cast_sparse(id, word)
# save ids and ratings
ids_titles <- as.integer(rownames(m_title))
ratings_titles <- df$rating[ids_titles]

ids_desc <- as.integer(rownames(m_desc))
ratings_desc <- df$rating[ids_desc]
# create new df with
df_word_matrix_title <-cbind(m_title, log_ratings = df$rating[ids_titles])
df_word_matrix_desc <-cbind(m_desc, log_ratings = df$rating[ids_desc])
# run glm on our data

library(doParallel)
cl <- makePSOCKcluster(CPU_CORES)
registerDoParallel(cl)
#cv_model <- cv.glmnet(m, df$price[ids_titles], parallel = TRUE)
cv_model_title <- cv.glmnet(m_title, ratings_titles, parallel = TRUE)
cv_model_desc <- cv.glmnet(m_desc, ratings_desc, parallel = TRUE)
stopCluster(cl)

```

```{r}

plot(cv_model_title)
plot(cv_model_desc)
# change 1se to red dashed
abline(v=log(cv_model$lambda.1se),col='red', lt=20)
#  build lexicons for sentiment - use the lambda obtained and get values of the  estimates for each word
lexicon_title <- cv_model_title$glmnet.fit %>%
  tidy() %>%
  filter(lambda == cv_model_title$lambda.1se,
         term != "(Intercept)",
         term != "log_ratings") %>%
  dplyr::select(word = term, coefficient = estimate)

lexicon_desc<- cv_model_desc$glmnet.fit %>%
  tidy() %>%
  filter(lambda == cv_model_desc$lambda.1se,
         term != "(Intercept)",
         term != "log_ratings") %>%
  dplyr::select(word = term, coefficient = estimate)


posnegwords_title <- lexicon_title %>%
    arrange(coefficient) %>%
    group_by(posneg = ifelse(coefficient < 0, "Negative", "Positive")) %>%
    top_n(30, abs(coefficient)) %>%
    ungroup()

posnegwords_desc <- lexicon_desc %>%
    arrange(coefficient) %>%
    group_by(posneg = ifelse(coefficient < 0, "Negative", "Positive")) %>%
    top_n(30, abs(coefficient)) %>%
    ungroup()


posnegwords_title%>%
    mutate(word = fct_reorder(word, coefficient)) %>%
    ggplot(aes(word, coefficient, fill = posneg)) +
    geom_col() +
    coord_flip() +
    labs(x = "",
         y = "Effect of word",
         title = "menuitem title +/- effect of words")


posnegwords_desc%>%
    mutate(word = fct_reorder(word, coefficient)) %>%
    ggplot(aes(word, coefficient, fill = posneg)) +
    geom_col() +
    coord_flip() +
    labs(x = "",
         y = "Effect of word",
         title = "menuitem description +/- effect of words")



# # visualize  positive wordcloud
# par(mfrow = c(1, 2))
# df4[df4$coefficient>0,] %>%
#   count(word) %>%
#   with(wordcloud(word, n, max.words = 100, colors = 'green'))
#
# #isualize  negative wordcloud
# df4[df4$coefficient<0,] %>%
#   count(word) %>%
#   with(wordcloud(word, n, max.words = 100, colors = 'red'))


```

```{r}
#df_title_merged<-left_join(df_title_agg,lexicon_title,by='word')
#df_desc_merged<-left_join(df_desc_agg,lexicon_desc,by='word')
# remove stopwords associated with wine that have little meaning
# paste0(sort(x),collapse = "','")
local_stopwords = c('–','and')
stopworddf<- data.frame(local_stopwords,"Custom")
colnames(stopworddf) = c("word","lexicon")
# add SMART lexicon
allstopwords <- rbind(stop_words,stopworddf)

# tokenized.df <- df %>%
#   dplyr::select( categories, title, description, price) %>%
#   unnest_tokens(word, description) %>%
#   anti_join(allstopwords) %>%
#   filter(!str_detect(word, "[0-9]"))


dfbigram_title <- df %>%
  dplyr::select(categories, title, description, price) %>%
  unnest_tokens(bigram, title, token = "ngrams", n = 2) %>% #bigram
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% allstopwords$word,
         !word2 %in% allstopwords$word,
         !str_detect(word1, "[0-9]"),
         !str_detect(word2, "[0-9]")) %>%
  unite(bigram, word1, word2, sep = " ")


dfbigram_desc <- df %>%
  dplyr::select(categories, title, description, price) %>%
  unnest_tokens(bigram, description, token = "ngrams", n = 2) %>% #bigram
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% allstopwords$word,
         !word2 %in% allstopwords$word,
         !str_detect(word1, "[0-9]"),
         !str_detect(word2, "[0-9]")) %>%
  unite(bigram, word1, word2, sep = " ")


dfbigram_title %>%
  count(bigram) %>%
  top_n(50,n) %>%
  arrange(desc(n)) %>%
  ggplot(aes(x = reorder(bigram, n), y = n)) +
  geom_bar(stat = "identity", color = "navyblue", fill = "royalblue") +
  theme(legend.position = "none") +
  coord_flip() +
  labs(y = "Frequency",
       x = "Top bigrams",
       title = "Top bigrams used",
       subtitle = "")





dfbigram_desc %>%
  count(bigram) %>%
  top_n(50,n) %>%
  arrange(desc(n)) %>%
  ggplot(aes(x = reorder(bigram, n), y = n)) +
  geom_bar(stat = "identity", color = "navyblue", fill = "royalblue") +
  theme(legend.position = "none") +
  coord_flip() +
  labs(y = "Frequency",
       x = "Top bigrams",
       title = "Top bigrams used",
       subtitle = "")





```

```{r}
# dfbigram %>%
#   filter(region %in% unname(unlist(top5_variety_region))) %>%
#   group_by(region, bigram) %>%
#   tally() %>%
#   top_n(10,n) %>%
#   arrange(desc(n)) %>%
#   ggplot(aes(x = reorder(bigram, n), y = n, fill = factor(region))) +
#   geom_bar(stat = "identity") +
#   theme(legend.position = "none") +
#   facet_wrap(~ region, scales = "free") +
#   coord_flip() +
#   labs(y = "Frequency",
#        x = "bigrams",
#        title = "bigrams used by top5 region",
#        subtitle = "")



# test/train split
dfnew2 <- df %>%  na.omit()  %>% sample_n(19000)
train_indexes <- createDataPartition(dfnew2$rating, times = 1, p = 0.7, list = FALSE)
train <- dfnew2[train_indexes,]
test  <- dfnew2[-train_indexes,]

train.tokens <- tokens(train$description, what = "word",
                       remove_numbers = TRUE, remove_punct = TRUE,
                       remove_symbols = TRUE, remove_hyphens = TRUE)
train.tokens <- tokens_tolower(train.tokens)
train.tokens <- tokens_select(train.tokens, allstopwords,
                              selection = "remove")
train.tokens <- tokens_wordstem(train.tokens, language = "english")
train.tokens.dfm <- dfm(train.tokens, tolower = FALSE)
train.tokens.dfm <- dfm_trim(train.tokens.dfm, sparsity = 0.99)
train.tokens.matrix <- as.matrix(train.tokens.dfm)
train.tokens.df <- cbind(Label = train$rating, convert(train.tokens.dfm, to = 'data.frame'))


# maybe use bm25
tf.idf <- function(x, idf) {
  x * idf
}
# term freq
termfreq <- apply(train.tokens.matrix, 1,  function(x) {
  x / sum(x)
})
# idf
idf <- apply(train.tokens.matrix, 2, function(x){
  corpus.size <- length(x)
  c <- length(which(x > 0))
  log(corpus.size / c)
})
# calculate tfidf
tfidf <- apply(termfreq, 2, tf.idf, idf = idf)
tfidf <- t(tfidf)
incomplete.cases <- which(!complete.cases(tfidf))
tfidf[incomplete.cases,] <- rep(0.0, ncol(tfidf))

# add features back
tfidf <- cbind(Label = train$rating, data.frame(tfidf), train$price)

# train model with linear SVM ane 10-fold Cross Validation
# may take forever to perform cross validation
library(doParallel)
cl <- makePSOCKcluster(CPU_CORES)
registerDoParallel(cl)
all_svm_token_models <- train(Label~., data = tfidf, method = 'svmLinear3', trControl = trainControl(method = "cv", number = 10, allowParallel=TRUE),parallel = TRUE)
plot(all_svm_token_models)
all_svm_token_models$finalModel
stopCluster(cl)

# todo: apply weighted cv_model lexicon sentitment ratings to tfidf to get the full effect

```


# Predictions/Recommendations

We can utilize a postive/negative word approach using `glmnet.fit` and create a ratings the words that are utilzized in descriptive text, for example in a unigram setting we see that `Mcdonalds menu items` actually has a negative coefficient to it.

```{r}
negative_words<-lexicon %>% arrange(coefficient)
positive_words<-lexicon %>% arrange(desc(coefficient))

positive_words[1:10,] %>%  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)

negative_words[1:10,] %>%  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```
